### 2024-07-02 paper/coverletter.tex
Fixes after SZ review

\documentclass[a4paper,12pt]{article}
\usepackage[a4paper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage[UKenglish]{isodate}

\linespread{1.5}
\setlength{\parindent}{0pt}

\begin{document}

% # Speech processing and communication systems

\begin{flushright}
  Paweł Antoniuk \\
  Faculty of Computer Science \\
  Białystok University of Science \\
  Wiejska 45A, 15-351 Białystok, Poland \\
  \href{mailto:pawel.antoniuk@sd.pb.edu.pl}{pawel.antoniuk@sd.pb.edu.pl} \\
  \vspace{0.5cm}
  \today
\end{flushright}

\vspace{1cm}

\noindent
Editorial Board \\
Archives of Acoustics

\vspace{1cm}

\noindent
Dear Editor,

We submit our manuscript ``Estimating Ensemble Location and Width in Binaural Recordings of Music with Convolutional Neural Networks'' for publication in Archives of Acoustics. Authored by Paweł Antoniuk (corresponding author) and Sławomir Krzysztof Zieliński, this paper presents a method for estimating spatial parameters in binaural audio recordings using a novel ensemble-based approach employing a spectrogram-based deep learning network technique.

\begin{itemize}
  \item We submit this paper to Archives of Acoustics as its content aligns with the journal's scope, particularly in \textbf{speech processing and communication systems}.
  \item We confirm this manuscript is original, unpublished, and not under consideration elsewhere.
  \item We recommend the following reviewers:
        \begin{itemize}
          \item \textbf{prof. dr hab. inż. Bożena Kostek} \\*
          Wydział Elektroniki, Telekomunikacji i Informatyki \\*
          Gdańsk University of Technology \\*
          ul. Gabriela Narutowicza 11/12 \\*
          80-233 Gdańsk \\*
          Poland \\*
          Email: \href{mailto:bozenka@sound.eti.pg.gda.pl}{bozenka@sound.eti.pg.gda.pl}
          \item \textbf{prof. dr hab. inż. Piotr Kleczkowski} \\*
          Wydział Inżynierii Mechanicznej i Robotyki \\*
          AGH University of Krakow \\*
          al. Mickiewicza 30 \\*
          30-059 Kraków \\*
          Poland \\*
          Email: \href{mailto:kleczkow@agh.edu.pl}{kleczkow@agh.edu.pl}
          \item \textbf{dr hab. inż. Grzegorz Szwoch} \\*
          Wydział Elektroniki, Telekomunikacji i Informatyki \\*
          Gdańsk University of Technology \\*
          ul. Gabriela Narutowicza 11/12 \\*
          80-233 Gdańsk \\*
          Poland \\*
          Email: \href{mailto:grzegorz.szwoch@pg.edu.pl}{grzegorz.szwoch@pg.edu.pl}
        \end{itemize}
  \item We propose classifying this manuscript as a \textbf{``research paper''} due to its empirical content.
  \item While the main content is within 17 pages, the manuscript exceeds the recommended 20-page limit to provide comprehensive context, detailed methodology, and extensive references. We believe this additional content is crucial for a full understanding of our research and to allow others to replicate our work.
\end{itemize}

We appreciate your consideration of our manuscript.

\vspace{1cm}

Sincerely, \\
Paweł Antoniuk

\end{document}